---
title: "PCA"
author: "rheaprag"
date: "2020-12-13"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## All Populations
Since the data was pruned in the previous step, new bfiles need to be created for the data:

```{}
./plink --bfile hapmap3_01_5 --extract hapmap3_allpop.prune.in --make-bed --out hapmap3_allpop_pruned
```

These new bfiles can be used to run the PCA with the population (not separated out!)
```{}
./plink --bfile hapmap3_allpop_pruned --pca 10 --out hapmap3_pca
```

To plot the data:
```{r, echo=FALSE}
library(ggplot2)
pcs <- read.table("hapmap3_pca.eigenvec",header=T)
eval <- scan("hapmap3_pca.eigenval")[1:10]
skree <- round(eval/sum(eval),3)
skree <- cbind.data.frame(skree,c(1,2,3,4,5,6,7,8,9,10))
colnames(skree) <- c("percent_var","PC")
ggplot(data=skree, aes(x=PC, y=percent_var)) + geom_point() + geom_line() + scale_x_continuous(breaks = 1:10) + ggtitle("Proportion of variance explained - Total Population")
```

Similar steps are going to be taken for the various population subsets:

## African

```{}
./plink --bfile hapmap_African --extract hapmap3_02_1_African.prune.in --make-bed --out hapmap3_African_pruned
```

```{}
./plink --bfile hapmap3_African_pruned --pca 10 --out hapmap3_African_pca
```

To plot the data:
```{r, echo=FALSE}
library(ggplot2)
pcs <- read.table("hapmap3_African_pca.eigenvec",header=T)
eval <- scan("hapmap3_African_pca.eigenval")[1:10]
skree <- round(eval/sum(eval),3)
skree <- cbind.data.frame(skree,c(1,2,3,4,5,6,7,8,9,10))
colnames(skree) <- c("percent_var","PC")
ggplot(data=skree, aes(x=PC, y=percent_var)) + geom_point() + geom_line() + scale_x_continuous(breaks = 1:10) + ggtitle("Proportion of variance explained - African")
```

## East Asia
```{}
./plink --bfile hapmap_EastAsia --extract hapmap3_02_1_EastAsia.prune.in --make-bed --out hapmap3_EastAsia_pruned
```

```{}
./plink --bfile hapmap3_EastAsia_pruned --pca 10 --out hapmap3_EastAsia_pca
```

To plot the data:
```{r, echo=FALSE}
library(ggplot2)
pcs <- read.table("hapmap3_EastAsia_pca.eigenvec",header=T)
eval <- scan("hapmap3_EastAsia_pca.eigenval")[1:10]
skree <- round(eval/sum(eval),3)
skree <- cbind.data.frame(skree,c(1,2,3,4,5,6,7,8,9,10))
colnames(skree) <- c("percent_var","PC")
ggplot(data=skree, aes(x=PC, y=percent_var)) + geom_point() + geom_line() + scale_x_continuous(breaks = 1:10) + ggtitle("Proportion of variance explained - East Asia")
```

## Utah
```{}
./plink --bfile hapmap_Utah --extract hapmap3_02_1_Utah.prune.in --make-bed --out hapmap3_Utah_pruned
```

```{}
./plink --bfile hapmap3_Utah_pruned --pca 10 --out hapmap3_Utah_pca
```

To plot the data:
```{r, echo=FALSE}
library(ggplot2)
pcs <- read.table("hapmap3_Utah_pca.eigenvec",header=T)
eval <- scan("hapmap3_Utah_pca.eigenval")[1:10]
skree <- round(eval/sum(eval),3)
skree <- cbind.data.frame(skree,c(1,2,3,4,5,6,7,8,9,10))
colnames(skree) <- c("percent_var","PC")
ggplot(data=skree, aes(x=PC, y=percent_var)) + geom_point() + geom_line() + scale_x_continuous(breaks = 1:10) + ggtitle("Proportion of variance explained - Utah")
```

## Colorado Chinese
```{}
./plink --bfile hapmap_ColoradoChinese --extract hapmap3_02_1_ColoradoChinese.prune.in --make-bed --out hapmap3_ColoradoChinese_pruned
```

```{}
./plink --bfile hapmap3_ColoradoChinese_pruned --pca 10 --out hapmap3_ColoradoChinese_pca
```

To plot the data:
```{r, echo=FALSE}
library(ggplot2)
pcs <- read.table("hapmap3_ColoradoChinese_pca.eigenvec",header=T)
eval <- scan("hapmap3_ColoradoChinese_pca.eigenval")[1:10]
skree <- round(eval/sum(eval),3)
skree <- cbind.data.frame(skree,c(1,2,3,4,5,6,7,8,9,10))
colnames(skree) <- c("percent_var","PC")
ggplot(data=skree, aes(x=PC, y=percent_var)) + geom_point() + geom_line() + scale_x_continuous(breaks = 1:10) + ggtitle("Proportion of variance explained - Colorado Chinese")
```

## Gujarati
```{}
./plink --bfile hapmap_Gujarati --extract hapmap3_02_1_Gujarati.prune.in --make-bed --out hapmap3_Gujarati_pruned
```

```{}
./plink --bfile hapmap3_Gujarati_pruned --pca 10 --out hapmap3_Gujarati_pca
```

To plot the data:
```{r, echo=FALSE}
library(ggplot2)
pcs <- read.table("hapmap3_Gujarati_pca.eigenvec",header=T)
eval <- scan("hapmap3_Gujarati_pca.eigenval")[1:10]
skree <- round(eval/sum(eval),3)
skree <- cbind.data.frame(skree,c(1,2,3,4,5,6,7,8,9,10))
colnames(skree) <- c("percent_var","PC")
ggplot(data=skree, aes(x=PC, y=percent_var)) + geom_point() + geom_line() + scale_x_continuous(breaks = 1:10) + ggtitle("Proportion of variance explained - Gujarati")
```

## Mexican
```{}
./plink --bfile hapmap_Mexican --extract hapmap3_02_1_Mexican.prune.in --make-bed --out hapmap3_Mexican_pruned
```

```{}
./plink --bfile hapmap3_Mexican_pruned --pca 10 --out hapmap3_Mexican_pca
```

To plot the data:
```{r, echo=FALSE}
library(ggplot2)
pcs <- read.table("hapmap3_Mexican_pca.eigenvec",header=T)
eval <- scan("hapmap3_Mexican_pca.eigenval")[1:10]
skree <- round(eval/sum(eval),3)
skree <- cbind.data.frame(skree,c(1,2,3,4,5,6,7,8,9,10))
colnames(skree) <- c("percent_var","PC")
ggplot(data=skree, aes(x=PC, y=percent_var)) + geom_point() + geom_line() + scale_x_continuous(breaks = 1:10) + ggtitle("Proportion of variance explained - Mexican")
```

## Toscans
```{}
./plink --bfile hapmap_Toscans --extract hapmap3_02_1_Toscans.prune.in --make-bed --out hapmap3_Toscans_pruned
```

```{}
./plink --bfile hapmap3_Toscans_pruned --pca 10 --out hapmap3_Toscans_pca
```

To plot the data:
```{r, echo=FALSE}
library(ggplot2)
pcs <- read.table("hapmap3_Toscans_pca.eigenvec",header=T)
eval <- scan("hapmap3_Toscans_pca.eigenval")[1:10]
skree <- round(eval/sum(eval),3)
skree <- cbind.data.frame(skree,c(1,2,3,4,5,6,7,8,9,10))
colnames(skree) <- c("percent_var","PC")
ggplot(data=skree, aes(x=PC, y=percent_var)) + geom_point() + geom_line() + scale_x_continuous(breaks = 1:10) + ggtitle("Proportion of variance explained - Toscans")
```
